Terrain is a stronger predictor of peat depth than airborne radiometrics in Norwegian landscapes

### 2025-08-05 ms/ms.Rmd
address TE.3

Although radiometric data must be in units of counts per second to model attenuation directly [@olearyDigitalSoilMapping2022], we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration has no effect on the tree-based machine learning algorithm that we used to model peat depth [@hastieElementsStatisticalLearning2009].

>>>>

Although radiometric data must be in units of counts per second to model attenuation directly [@olearyDigitalSoilMapping2022], we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration has no effect on the tree-based machine learning algorithm that we used to model peat depth [@hastieElementsStatisticalLearning2009].
We also used the total count variable as provided, rather than calculating a gamma dose rate based on the potassium, thorium, and uranium [as was done in @gatisMappingUplandPeat2019], because these were highly correlated at both study sites ($\rho = 0.989, 0.986$), and because conversions to dose rates are approximations [@iaeaGuidelinesRadioelementMapping2003].

### 2025-08-05 ms/ms.tex
address TE.3

Although radiometric data must be in units of counts per second to model attenuation directly \citep{olearyDigitalSoilMapping2022}, we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration has no effect on the tree-based machine learning algorithm that we used to model peat depth \citep{hastieElementsStatisticalLearning2009}.

>>>>

Although radiometric data must be in units of counts per second to model attenuation directly \citep{olearyDigitalSoilMapping2022}, we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration has no effect on the tree-based machine learning algorithm that we used to model peat depth \citep{hastieElementsStatisticalLearning2009}.
We also used the total count variable as provided, rather than calculating a gamma dose rate based on the potassium, thorium, and uranium \citep[as was done in][]{gatisMappingUplandPeat2019}, because these were highly correlated at both study sites (\(\rho = 0.989, 0.986\)), and because conversions to dose rates are approximations \citep{iaeaGuidelinesRadioelementMapping2003}.
